---
title: "Comparing Ireland to other countries"
author: "Catherine Hurley"
date: "3/23/2020"
output: 
  html_document:
    
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width=5, fig.height=5,message=FALSE, warning=FALSE)
```

```{r echo=FALSE}
suppressMessages(library(tidyverse))
suppressMessages(library(ggplot2))
suppressMessages(library(gridExtra))
suppressMessages(library(plotly))
```

#### Get the data

```{r echo=FALSE}
devtools::install_github("RamiKrispin/coronavirus")

```
Use the data from coronavirus, which is from Johns Hopkins.
```{r}
devtools::install_github("RamiKrispin/coronavirus") # for up to date version

library(coronavirus)
```

#### Set up a daily dataset

```{r}
daily <- coronavirus%>% mutate(deaths= ifelse(type=="death", cases, 0)) %>%
  group_by(Country.Region, date) %>%
  summarise(cases=sum(cases), deaths=sum(deaths))%>%
  ungroup()%>%
  group_by(Country.Region) %>%
  mutate(totcases=cumsum(cases), totdeaths=cumsum(deaths)) %>%
  ungroup()%>%
  filter(totcases >0)
```

#### To join in populations..

```{r}

if (!require("wbstats")) install.packages("wbstats")
library(wbstats)

country_data <- 
  left_join(wb(indicator = c("SP.POP.TOTL", "EN.POP.DNST","SP.URB.TOTL.IN.ZS", "UNDP.HDI.XD"),country="countries_only" ,
               mrv=1,return_wide = TRUE),
          wbcountries()) %>%
  select(country, region, pop=SP.POP.TOTL, popdensity=EN.POP.DNST, percenturban= SP.URB.TOTL.IN.ZS, hdi= SP.URB.TOTL.IN.ZS,income=income)

daily <- left_join(daily, country_data, by=c("Country.Region" ="country"))
```

To select other indices from wbstats, use wbsearch and pick the index you want.

```{r eval=FALSE}
View(wbsearch("population"))
```


#### Countries similar to Ireland in population

```{r}
selcountries <- c("Ireland","Finland", "Denmark", "Norway") # similar populations
```

or, using pop_data

```{r}
ipop <- filter(country_data, country=="Ireland") 
selcountries <- filter(country_data, pop <= ipop$pop/0.8 & pop >= ipop$pop*.8 & region==ipop$region )$country
daily1 <- filter(daily, Country.Region %in% selcountries)

```





#### Plots of selected countries

```{r,fig.width=6, fig.height=3}
# by date----------- totcases

# Use Ireland's first case
mindate <- filter(daily1, cases > 0 & Country.Region=="Ireland") %>% summarise(date=min(date))  %>% pull(date)

maxdate <- Sys.Date()

p1 <- ggplotly(ggplot(data=filter(daily1, date >= mindate),
       aes(x=date, y=totcases, colour=Country.Region))+ geom_line()+
  scale_y_continuous(trans='log10') +
  geom_text(data=filter(daily1, date == maxdate-4), aes(label=Country.Region), size=3)+ 
  theme(legend.position='none'))


# by date----------- deaths
p2<- ggplotly(ggplot(data=filter(daily1, date >= mindate),
       aes(x=date, y=totdeaths, colour=Country.Region))+ geom_line()+
  geom_text(data=filter(daily1, date == maxdate-4), aes(label=Country.Region), size=3)+ 
  theme(legend.position='none'))


subplot(p1,p2)

```


#### Line up the curves, here by second case

```{r,fig.width=6, fig.height=3}

ncases <- 2
daysn <- daily1%>% group_by(Country.Region) %>%
  filter(totcases>=ncases) %>% mutate(dayn =min(date)) %>% 
  select(Country.Region, dayn) %>%slice(1) 

daily1 <- left_join(daily1, daysn, by="Country.Region") %>%
  mutate(days = as.numeric(date-dayn))

# by days----------- totcases

p1<-ggplotly(ggplot(data=filter(daily1, days >=0),
       aes(x=days, y=totcases, colour=Country.Region))+ geom_line()+
  scale_y_continuous(trans='log10') +
  geom_text(data=filter(daily1, date == maxdate-4), aes(label=Country.Region), size=3)+ 
  theme(legend.position='none'))


# by days----------- deaths

p2 <- ggplotly(ggplot(data=filter(daily1, days >=0),
       aes(x=days, y=totdeaths, colour=Country.Region))+ geom_line()+
  geom_text(data=filter(daily1, date == maxdate-4), aes(label=Country.Region), size=3)+ 
  theme(legend.position='none'))

subplot(p1,p2)

```


#### Facet by percent urban popultion

```{r fig.width=8, fig.height=5, warnings=FALSE}

europe <- filter(daily, region=="Europe & Central Asia")
groupvar <-europe %>% group_by(Country.Region) %>% slice(1)%>% pull(percenturban)
facbreaks <- c(0,quantile(groupvar, (1:6)/6,na.rm=T))

europe$percenturbanG <- cut(europe$percenturban,breaks=facbreaks)
mindate <- filter(europe, totcases > 0 ) %>% summarise(date=min(date))  %>% pull(date)

# by date----------- cases
ggplotly(ggplot(data=filter(europe, !is.na(percenturban) & date >=mindate),
       aes(x=date, y=totcases, colour=Country.Region))+ geom_line()+
  scale_y_continuous(trans='log10') + facet_wrap(~percenturbanG)  +theme(legend.position='none'))

# by date----------- deaths

mindate1 <- filter(europe, totdeaths > 1 ) %>% summarise(date=min(date))  %>% pull(date)
ggplotly(ggplot(data=filter(europe, !is.na(percenturban) & date >=mindate1 ),
       aes(x=date, y=totdeaths, colour=Country.Region))+ geom_line()+
  scale_y_continuous(trans='log10') + facet_wrap(~percenturbanG)  +theme(legend.position='none'))

```

#### Facet by HDI

```{r fig.width=8, fig.height=5, warnings=FALSE}

europe <- filter(daily, region=="Europe & Central Asia")
groupvar <-europe %>% group_by(Country.Region) %>% slice(1)%>% pull(hdi)
facbreaks <- c(0,quantile(groupvar, (1:6)/6,na.rm=T))

europe$hdiG <- cut(europe$percenturban,breaks=facbreaks)
mindate <- filter(europe, totcases > 0 ) %>% summarise(date=min(date))  %>% pull(date)

# by date----------- cases
ggplotly(ggplot(data=filter(europe, !is.na(percenturban) & date >=mindate),
       aes(x=date, y=totcases, colour=Country.Region))+ geom_line()+
  scale_y_continuous(trans='log10') + facet_wrap(~hdiG)  +theme(legend.position='none'))

# by date----------- deaths

mindate1 <- filter(europe, totdeaths > 1 ) %>% summarise(date=min(date))  %>% pull(date)
ggplotly(ggplot(data=filter(europe, !is.na(percenturban) & date >=mindate1 ),
       aes(x=date, y=totdeaths, colour=Country.Region))+ geom_line()+
  scale_y_continuous(trans='log10') + facet_wrap(~hdiG)  +theme(legend.position='none'))

```

#### Facet by income

Not so useful becuase almost all European countries are high income

```{r fig.width=8, fig.height=2.5, warnings=FALSE,  eval=FALSE}

europe <- filter(daily, region=="Europe & Central Asia")

# by date----------- cases
ggplotly(ggplot(data=filter(europe, !is.na(percenturban) & date >=mindate),
       aes(x=date, y=totcases, colour=Country.Region))+ geom_line()+
  scale_y_continuous(trans='log10') + facet_wrap(~income)  +theme(legend.position='none'))

# by date----------- deaths

mindate1 <- filter(europe, totdeaths > 1 ) %>% summarise(date=min(date))  %>% pull(date)
ggplotly(ggplot(data=filter(europe, !is.na(percenturban) & date >=mindate1 ),
       aes(x=date, y=totdeaths, colour=Country.Region))+ geom_line()+
  scale_y_continuous(trans='log10') + facet_wrap(~income)  +theme(legend.position='none'))

```
